---
title: Weather in Berlin
format: 
  dashboard:
    orientation: columns
    nav-buttons: [github]
    github: https://github.com/mrguyperson/berlin_weather/
    theme: flatly
logo: images/berlin.png

---

```{r}
#| label: load-packages
library(tidyverse)
library(targets)
library(showtext)
library(glue)
library(openmeteo)
library(scales)
library(gt)
library(here)
```


```{r}
#| label: prep-data

source(here("R/functions.R"))

# for plot
tar_load(history_with_calendar)
tar_load(this_year)
tar_load(new_records)
tar_load(city)
tar_load(start_date)
tar_load(today)

# for value boxes
tar_load(hottest_year)
tar_load(coldest_year)
tar_load(temperature_slope)

# for map label
tar_load(lat)
tar_load(long)
tar_load(country)

# for tables
tar_load(top_10_hottest)
tar_load(top_10_coldest)

```

## Colum {width="80%"}

### Row - value boxes {height="20%"}
::: {.valuebox icon="fire" color="#d7191c"}
**Hottest year:**

`r hottest_year$year`

:::

::: {.valuebox icon="thermometer-snow" color="#2c7bb6"}
**Coldest year:**

`r coldest_year$year`

:::

::: {.valuebox icon= "exclamation-diamond" color="#daa520"}
**Annual temp. change since 1940:**

`r glue("{if_else(temperature_slope > 0, '+', '\u2212')}{temperature_slope}\u00B0C")`

:::

### Row - Plot {height="80%"}

::: {.card title="**Daily temperatures in Berlin since 1 Jan, 1940**"}


```{r}
#| label: average_daily_temp_fig
#| fig-width: 8
#| fig-asp: 0.35
make_plot(history_with_calendar, this_year, new_records, city, start_date, today)
```
Data provided by [Open-meteo](https://open-meteo.com/).
Data from leap days (i.e., 29 February) are excluded.
Last updated **`r make_nice_date(today())`**.

:::

## Column

### Row - map
``` {r}
#| label: interactive-map

library(leaflet)



cat("title=", glue("**{city}, {country} ({round(lat, 2)}\u00B0, {round(long,2)}\u00B0)**"))

leaflet() |> addTiles() |> 
  setView(long, lat, zoom = 9)
```

### Row - temp tables {.tabset height="60%"}

::: {.card title="10 hottest"}
**Most extreme temperatures in dataset**
``` {r}
#| label: hottest_temps

top_10_hottest %>%
    mutate(date = make_nice_date(date)) %>%
    rename(Date = date, "Temperature (\u00B0C)" := temperature) %>%
    gt() %>%
    cols_align(align = "right")# %>%
    # data_color(
    #     method = "numeric",
    #     palette = "nord::victory_bonds"
    # )


```

:::

::: {.card title="10 coldest"}
**Most extreme temperatures in dataset**
```{r}

top_10_coldest %>%
    mutate(date = make_nice_date(date)) %>%
    rename(Date = date, "Temperature (\u00B0C)" := temperature) %>%
    gt() %>%
    cols_align(align = "right")

```

:::